---
title: Long-term nutrient enrichment alters coastal plain wetland plant-microbe relationships
author: "Daniya Stephens, Aied Garcia, Jasper Leavitt, Regina Bledsoe, Carol Goodwillie, Ariane L. Peralta"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  fig_caption: yes
  pdf_document: null
header-includes:
- \usepackage{array}
- \usepackage{graphics}
- \usepackage[utf8]{inputenc}
---

Project Description: Fill out

# Initial Setup
```{r Initial Setup, include=FALSE}
rm(list=ls())
setwd("~/GitHub/WRC2014to2020_PlantMicro/analyses")
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
ci <- function(x, ...){1.96 * sd(x,na.rm = TRUE)}

# Code Dependencies
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")
require("vegan"); require("lme4"); require("reshape"); require("ggplot2")
require("ade4"); require("png"); require ("agricolae"); require("tidyr")
require("dplyr"); require("reshape2"); require("picante"); require("phyloseq")
```
#Import Files
## Environmental Data
```{r Import Files - Env, include=FALSE}
# Import Environmental Data
#design <- read.csv("..data/WRC_2014to2020_design.csv") #2014-2020 data
design <- read.csv("../data/WRC_2014_2015_2018_design_env.csv", row.names=1) 
```
## Bacterial Data - including bulk and tea
```{r Import Files - Bacteria, include=FALSE}
# Import OTU data
# Import Raw Data
otu.in <- read.otu("../data/WRC141518.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.shared")

# Correct Sample IDs and Subset File
missing <- setdiff(rownames(otu.in), rownames(design.2018))
otu <- otu.in[-(which(rownames(otu.in) %in% missing)), ]
dim(otu)
otu <- otu[match(rownames(design.2018), rownames(otu)),]
all.equal(rownames(otu), rownames(design.2018))
dim(otu)
# OTU table - remove otus w/ < 2 occurrences across all sites
otus <- otu[, which(colSums(otu) >= 2)]
dim(otus)

# Make Presence Absence Matrix
dataPA <- (otus > 0) * 1

# Make Relative Abundance Matrices
dataREL <- otus
for(i in 1:dim(otus)[1]){
  dataREL[i,] <- otus[i,]/sum(otus[i,])
}

# Import Taxonomy File
otu.tax <- read.tax(taxonomy = "../data/WRC141518.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.opti_mcc.0.03.cons.taxonomy",
                   format = "rdp", tax.levels = 6, col.tax = 3)
```

## Subset for Rhizo
```{r}
rhizo <- otu.tax$OTU[which(otu.tax$Order == "Rhizobiales")]

rhizo.otu <- otu[, which(colnames(otu) %in% rhizo)]
rhizo.tax <- otu.tax[which(otu.tax$OTU %in% rhizo), ]
```

# Diversity Metrics - Hypothesis Testing
```{r Diversity Metrics - Bacteria, echo=TRUE}
# Rarefy Abundances (min abundance is 29392. We are sampling to 29000)
min(rowSums(otus))
otus.r <- rrarefy(otus, 29000)

# Fisher's Alpha
fisher <- fisher.alpha(otus.r)

# Species Richness
#richness <- rowSums((PWESdata.r >= 1))
richness <- rowSums((otus >= 1))

# Shannon Diversity
shannon <- diversity(otus.r, "shannon")

# Simpson's Evenness
simp.even <- apply(otus.r, 1, simp_even)

#Pielou’s evenness
J <- shannon/log(specnumber(otus.r[,-c(1:1)]))

#combined richness, diversity, evenness
diversity <- cbind(design.2018,richness,shannon,simp.even,J)
write.csv(diversity,"../data/2018.diversity.bact.raw.csv")
```

# Diversity Metrics - Hypothesis Testing
```{r Hypothesis Testing - Bacteria, echo=TRUE}
library(lmerTest)
diversity$ditch <- as.factor(diversity$ditch)
str(diversity)

shannon.lm <- lmer(shannon ~ treatment*ditch + (1|block), data = diversity)
plot(shannon.lm)
shannon.lm
anova(shannon.lm, type="II", ddf="Kenward-Roger")
```

#Plot shannon diversity 
```{r Plot - Shannon Diversity, echo=TRUE}
ditch.labs <- c("wet_0ditch", "dry_1ditch") #for facet labels
names(ditch.labs) <- c("0", "1")

# Graphing Shannon Diversity
p <- ggplot(diversity, aes(x=treatment, y=shannon, color=as.factor(treatment)))+ geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("mowed", "mowed/fertilized")) 
p1=p+geom_smooth(method="lm")+facet_wrap(~ditch)+facet_grid(. ~ ditch,labeller = labeller(ditch=ditch.labs))
shannon<-p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Treatment", y = "Shannon Diversity Index (H')") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("M", "MF"), labels=c("mowed", "mowed/fertilized"))
shannon
ggsave("../figures/2018.bacteria.shannon.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
# Simple Hypothesis Testing - Microbes
```{r perMANOVA - Bacteria - All sources, echo=TRUE}
#PERMANOVA
new.data <-cbind(design.2018,dataREL)
adonis = adonis(new.data[,-c(1:16)]~treatment*ditch, method = "bray", data = new.data, perm=1000)
adonis
```
# Bacterial Ordinations
```{r Ordination (PCoA) - Bacteria - bulk and teas, include=FALSE}
# Principal Coordinates Analysis
dataREL.dist <- vegdist(dataREL, method="bray")

pcoa <- cmdscale(dataREL.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1b <- round(pcoa$eig[1] / sum(pcoa$eig), 3) * 100
explainvar2b <- round(pcoa$eig[2] / sum(pcoa$eig), 3) * 100
sum.eigb <- sum(explainvar1b, explainvar2b)

explainvar1b #26.6
explainvar2b #18.1
```

```{r Ordination (PCoA) - Bacteria, include=FALSE}
# Principal Coordinates Analysis - bulk and teas
#new.data <-cbind(design_final,dataREL)
pcoa.groups <- paste(new.data$treatment, new.data$ditch, sep = "_")
pcoa.points <- data.frame(pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)
trt <- c("M","M","MF","MF")
ditch <- c("wet_0ditch","dry_1ditch","wet_0ditch","dry_1ditch")
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts$trt <- as.factor(trt)
pcoa.cent.dataframe.trts$ditch <- as.factor(ditch)

#Plot
df1a <- as.data.frame(pcoa.cent.dataframe.trts)
plot1a <- ggplot(df1a, aes(x=V1, y=V2, colour=trt, shape = ditch)) + theme_bw() 
plot1a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_point(aes(colour=trt), size=5, stroke = 1.25, show.legend = TRUE) +
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") + 
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
  scale_colour_manual(labels = c("mowed","mowed/fertilized"), 
                      values = c("gray70", "darkgreen")) +
  scale_shape_manual(labels = c("(wet_0ditch)","(dry_1ditch)"), values = c(15,22)) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (26.6%)") + ylab("PCoA 2 (18.1%)") +
        labs(colour = "Treatment", shape = "Ditch") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4)),
           shape = guide_legend(override.aes = list(size = 4))) +
    ggtitle("Bacterial Community Composition (2018)")
ggsave("../figures/2018.bact.ordination.bulk.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
## Plant Data - 2018 only
```{r Import Files - Plant, include=FALSE}
PCC <- read.csv("../data/WRC18_plant_RA.csv", header=TRUE)
PCC.2018 <- subset(PCC, treatment == "M" | treatment == "MF")
PCC.2018 <- PCC.2018[,-c(1,2,84)]
labels(PCC.2018)
PCC.2018.a <- PCC.2018[, which(colSums(PCC.2018) > 0)]
PCC.2018 <- PCC.2018.a
```
# Diversity Metrics - Hypothesis Testing
```{r Diversity Metrics - Plant, echo=TRUE}
# Fisher's Alpha
PCC.2018.PA <- decostand(PCC.2018, method="pa", na.rm=TRUE)
fisher <- fisher.alpha(PCC.2018.PA)

# Species Richness
richness <- rowSums((PCC.2018.PA >= 1))

# Shannon Diversity
shannon <- diversity(PCC.2018, "shannon")

# Simpson's Evenness
simp.even <- apply(PCC.2018, 1, simp_even)

#Pielou’s evenness
J <- shannon/log(specnumber(PCC.2018[,-c(1:1)]))

#combined richness, diversity, evenness
diversity.plant <- cbind(design.2018,richness,shannon,simp.even,J)
write.csv(diversity.plant,"../data/2018.diversity.plant.csv")
```

# Diversity Metrics - Hypothesis Testing
```{r Hypothesis Testing - Plant, echo=TRUE}
## INCOMPLETE
library(lmerTest)
diversity$ditch <- as.factor(diversity.plant$ditch)
str(diversity)

shannon.lm <- lmer(shannon ~ treatment*ditch + (1|block), data = diversity.plant)
plot(shannon.lm)
shannon.lm
anova(shannon.lm, type="II", ddf="Kenward-Roger")
```

#Plot shannon diversity 
```{r Plot - Plant Shannon Diversity, echo=TRUE}
ditch.labs <- c("wet_0ditch", "dry_1ditch") #for facet labels
names(ditch.labs) <- c("0", "1")

# Graphing Shannon Diversity
p <- ggplot(diversity.plant, aes(x=treatment, y=shannon, color=as.factor(treatment)))+ geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("mowed", "mowed/fertilized")) 
p1=p+geom_smooth(method="lm")+facet_wrap(~ditch)+facet_grid(. ~ ditch,labeller = labeller(ditch=ditch.labs))
shannon.plant <-p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Treatment", y = "Shannon Diversity Index (H')") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("M", "MF"), labels=c("mowed", "mowed/fertilized"))
shannon.plant
ggsave("../figures/2018.plant.shannon.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
# Simple Hypothesis Testing - Microbes
```{r perMANOVA - Plant, echo=TRUE}
#PERMANOVA
new.data <-cbind(design.2018,PCC.2018)
adonis = adonis(new.data[,-c(1:16)]~treatment*ditch, method = "bray", data = new.data, perm=1000)
adonis
```
# Plant Ordinations
```{r Ordination (PCoA) - Bacteria - bulk and teas, include=FALSE}
# Principal Coordinates Analysis
dataREL.dist <- vegdist(PCC.2018, method="bray")

pcoa <- cmdscale(dataREL.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1b <- round(pcoa$eig[1] / sum(pcoa$eig), 3) * 100
explainvar2b <- round(pcoa$eig[2] / sum(pcoa$eig), 3) * 100
sum.eigb <- sum(explainvar1b, explainvar2b)

explainvar1b #38.0
explainvar2b #22.2
```

```{r Ordination (PCoA) - Bacteria, include=FALSE}
# Principal Coordinates Analysis - plants
pcoa.groups <- paste(new.data$treatment, new.data$ditch, sep = "_")
pcoa.points <- data.frame(pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)
trt <- c("M","M","MF","MF")
ditch <- c("wet_0ditch","dry_1ditch","wet_0ditch","dry_1ditch")
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts$trt <- as.factor(trt)
pcoa.cent.dataframe.trts$ditch <- as.factor(ditch)

#Plot
df2a <- as.data.frame(pcoa.cent.dataframe.trts)
plot2a <- ggplot(df2a, aes(x=V1, y=V2, colour=trt, shape = ditch)) + theme_bw() 
plot2a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_point(aes(colour=trt), size=5, stroke = 1.25, show.legend = TRUE) +
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") + 
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
  scale_colour_manual(labels = c("mowed","mowed/fertilized"), 
                      values = c("gray70", "darkgreen")) +
  scale_shape_manual(labels = c("(wet_0ditch)","(dry_1ditch)"), values = c(15,22)) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (38.0%)") + ylab("PCoA 2 (22.2%)") +
        labs(colour = "Treatment", shape = "Ditch") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4)),
           shape = guide_legend(override.aes = list(size = 4)))+
  ggtitle("Plant Community Composition (2018)")
ggsave("../figures/2018.plant.ordination.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
# CORRELATIONS Plant - Bacteria
```{r mantel test, include=FALSE}
#treatment and ditch included 
dist.plant <- vegdist(PCC.2018, method="bray")
dist.bact <- vegdist(dataREL, method="bray")

#separate by treatment
new.data <-cbind(design.2018,dataREL)
new.data.bact.M <- subset(new.data, treatment == "M")
new.data.bact.MF <- subset(new.data, treatment == "MF")
dist.bact.M <- vegdist(new.data.bact.M[,-c(1:16)], method="bray")
dist.bact.MF <- vegdist(new.data.bact.MF[,-c(1:16)], method="bray")

new.data.plant <-cbind(design.2018,PCC.2018)
new.data.plant.M <- subset(new.data.plant, treatment == "M")
new.data.plant.MF <- subset(new.data.plant, treatment == "MF")
dist.plant.M <- vegdist(new.data.plant.M[,-c(1:16)], method="bray")
dist.plant.MF <- vegdist(new.data.plant.MF[,-c(1:16)], method="bray")

# MANTEL TESTS
mantel.rtest(dist.plant, dist.bact, nrepet = 999) #all treatments and ditch included - Mantel r = 0.5193241, p = 0.001
mantel.rtest(dist.plant.M, dist.bact.M, nrepet = 999) # mowed only: Mantel r = 0.4825625, p = 0.014
mantel.rtest(dist.plant.MF, dist.bact.MF, nrepet = 999) # mowed/fert: Mantel r = 0.4083636, p = 0.036 

# correlation alpha diversity
alpha.plant.M <- subset(diversity.plant, treatment == "M")
alpha.plant.MF <- subset(diversity.plant, treatment == "MF")
alpha.bact.M <- subset(diversity, treatment == "M")
alpha.bact.MF <- subset(diversity, treatment == "MF")

cor.test(diversity.plant$shannon, diversity$shannon) # ALL: Pearsonn rho = -0.751, p = 0.0008
cor.test(alpha.plant.M$shannon, alpha.bact.M$shannon) # mowed only: Pearson rho = -0.6412016, p = 0.087
cor.test(alpha.plant.MF$shannon, alpha.bact.MF$shannon) # mowed/fert: Pearson rho = -0.4943074, p = 0.2131
plant.bact.shannon <- cbind(design.2018, diversity.plant$shannon, diversity$shannon)

p <- ggplot(plant.bact.shannon, aes(x=diversity.plant$shannon, y=diversity$shannon, color=treatment)) +
  scale_color_manual(name="Treatment", values=c("gray70","darkgreen"), labels = c("mowed", "mowed/fertilized")) +  
  geom_point(size=4)
p1=p+geom_smooth(method="lm")
shannon.cor <- p1 + theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
  theme(axis.title=element_text(vjust=1,size=16), 
        axis.text=element_text(size=16), 
        axis.text.x = element_text(vjust=0.65, hjust=0.5, size=16), 
        panel.border = element_rect(colour = "black",size=1)) +
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  labs(x = "Plant Shannon Diversity Index (H')", y = "Bacterial Shannon Diversity Index (H')") +
  theme(strip.text.x = element_text(size=16, face="bold"), strip.text.y = element_text(size=16, face="bold"), strip.background = element_rect(colour="black", fill="white", size=1))
shannon.cor
ggsave("../figures/2018.plant.bact.shannon.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```